Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass

### 2024-02-22 manuscript.Rmd
add script for table of species correspondance to tetum names

# References

>>>>

## Supplemental information

### Botanical identification from Tetum names

```{r tetum-names}
sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
sheets2 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)

vern1 <- lapply(sheets1, \(x){
  df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x) 
  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |> 
    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
})
vern2 <- lapply(sheets2, \(x){
    df = readxl::read_excel("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = x) 
  df = df[, c("TETUM_NAME", "SCIENTIFIC_NAME")] |> 
    unique() |>  subset(!SCIENTIFIC_NAME %in% c("Unknown", "NR", "Unkown"))
})

vern = rbind(rbindlist(vern1), rbindlist(vern2)) |>  unique()
vern <- subset(vern, !is.na(TETUM_NAME) & !is.na(SCIENTIFIC_NAME))

## correct scientific names 
vern[, c("genus", "species") := tstrsplit(SCIENTIFIC_NAME, " ")[1:2]]
vern[species %in% c("sp", "sp."), species := NA]
vern[species == "purpuroides", species := NA]
corr = BIOMASS::correctTaxo(genus = vern$genus, species = vern$species)
vern = cbind(vern, corr)

vern[, latin := paste(genusCorrected, speciesCorrected)]
vern[, latin := gsub(" NA", " sp.", latin)]
vern[,tetum := gsub("  ", " ", tolower(TETUM_NAME))]
vern <- vern[, c("tetum", "latin", "genusCorrected", "speciesCorrected")]
vern = unique(vern)

# check that all names are unique
# if one tetum name has two levels of id (same genus), keep better identification
dupl_tetum <- vern[duplicated(tetum), unique(tetum)]
vern = vern[!(tetum %in% dupl_tetum & is.na(speciesCorrected))]
setorder(vern, tetum)

vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
write.csv(vern, "data/species-correspondance.csv", row.names = FALSE) 
```
# References
